Bit Complexity for Computation of one Point in each Connected Component of a Smooth Real Hypersurface

### 2020-02-01 main.tex
Update main.tex
Updates from Overleaf

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
%\title{Bit Complexity for Computation of one Point in each Connected
%Component of a Smooth Real Hypersurface}
\title{On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  We present a full analysis of an efficient algorithm for the
  computation of at least one point in each connected component of a
  smooth real hypersurface, taking into account bit complexity and
  all coefficient growth.  This problem is a basic and important
  subroutine used in semi-algebraic geometry. For instance, it is used
  in determining an upper bound on the number of connected components
  of a real hypersurface. The algorithm requires some genericity
  properties, which we ensure through random changes of variables that
  produce generic coordinate systems. In order to ensure these
  properties with arbitrarily high probability, we developed a new
  quantitative extension of Thom's weak transversality theorem, and a
  new quantitative Noether normalization statement for polar
  varieties. These statements enable us to give a precise probability
  analysis of a monte carlo algorithm, where we can guarantee
  correctness with arbitrarily high probability.

  Algorithms for deciding connectivity queries in real algebraic
  geometry have been developed in a series of papers
  \cite{a,b,c,d}. However, these algorithms are based in the
  arithmetic complexity model, and analyses of the bit complexity
  questions were left for future work.  Our results are an important
  first step in this analysis.

  % Furthermore, as bit complexity provides a far more realistic model
  % of computation compared to arithmetic complexity, which only
  % counts each operation at unit cost whereas bit complexity provide
  % more information for both time and space resources that are needed
  % in practice, it is important that this missing gap is filled.
  % As bit complexity provides a far more realistic model of
  % computation compared to arithmetic complexity, which only counts
  % each operation at unit cost, our contribution is necessary and
  % important. Indeed, bit estimates provide more information for both
  % time and space resources that are needed in practice.
\end{abstract}

@@

\nl Use \cite{AD} to compute the partial derivatives $\frac{\pa f^A}{\Pr X_{i+1}},\hdots,\frac{\pa f^A}{\Pr X_n}$\;

@@

Let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A)[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ Recall from Section 4.1.4 that $f^\A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Throughout this section, for simplicity let $\I^\A$ denote $\I(\pi_i,V^\A)$ and let $W^\A$ denote $W(\pi_i,V^\A).$ To increase readability and simplify notation, we often do not specify $i$, as it is fixed.   

@@

Let $f \in \mathbb{Q}[X_1,\hdots,X_n]$ be squarefree with total degree $D$ and $V(f)$ smooth. For $i \in \{1,\hdots,n\}$, recall $\textbf{H}_i^{'}(4):$ \textit{$W(\pi_i,V)$ is in Noether position for $\pi_{i-1}$}, which is equivalent to having the extension  

@@

@@

then $A$ cancels none of the denominators of $a_{j}$ or $b_{j,l}$. Define 

@@

\Xi_j:=V(h_j), 

@@

where $h_j$ is the product of the denominators of the coefficients of $a_j$ and $b_{j,l}$, so that when $A \not \in \Xi_j$ then $A$ cancels none of the denominators $a_{j}$ or $b_{j,l}$ which provide the integral dependence relation. For each $j$, we want a bound on the degree of $\Xi_j$. We can obtain bounds by applying an effective Nullstellensatz theorem. However, the theorem requires $P_{j}$ to be a polynomial. We can obtain a polynomial from $P_{j}$ by clearing denominators. We can multiply $P_{j}$ by $d_j \in \Q[\A_{i,j}]$ equal to the least common multiple of the denominators of $P_{j}.$ Put 

@@

\overline{P_{j}} := dP_{j}\in \Q[\A][\textit{\textbf{X}}],

@@

$\deg \pjb\leq (D+1)^{n-i+1}.$ 

@@

\deg d \leq \deg \pjb \leq (D+1)^{n-i+1}.

@@

\alpha_i = \sum_{l=1}^{n-i+1} A_l(\A_{i,j},\textit{\textbf{X}},1/\pjb)g_l + 0

@@

A_l(\A_{i,j},\textit{\textbf{X}},1/\pjb) &= \sum_{i=1}^{n_l} A_{l,i}(\A_{i,j},\Xb,1/\pjb)\\
&= \sum_{i=1}^{n_l} \left(\frac{1}{P_j^{k_l}}\right)\At_{l,i}(\A_{i,j},\textit{\textbf{X}}), ~k_l \in \mathbb{Z}.

@@

Now let $k := \max\{k_l\}$ so that 

@@

    \pjb^k \alpha &= \pjb^k\left(\sum_{l=1}^{n-i+1} A_l(\A_{i,j},\textit{\textbf{X}},1/\pjb)g_l\right)\\
    &= \sum_{l=1}^{n-i+1}\left( \sum_{i=1}^{n_l}\left(\pjb^{k-k_{n_l}}\right)\At_l(\A_{i,j},\Xb) \right)g_l \in \Q[\A_{i,j}][\Xb],

@@

which shows that $\alpha$ is the only remaining denominator. We know that $\alpha$ is not zero, we know the leading coefficient of $\overline{P_j}$ is not zero, we have an upper bound for the degree of $\alpha$, and we know $\deg d \leq \deg P_j = (D+1)^{n-i+1}$, where $\pjb=dP_j.$ By making $\pjb$ monic, the integrality of 
\begin{align*}\mathbb{Q}(\frak{A}_{i,j})[\textbf{X}_{\leq i-1}]\rightarrow\mathbb{Q}(\frak{A}_{i,j})[\textbf{X}]/\I_i^{\A}

@@

holds with $\pjb(X_j)\in \I_i^{\A}.$

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Bit Complexity for Computation of one Point in each Connected Component of a Smooth Real Hypersurface}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
 We present a full analysis of an efficient algorithm for the
 computation of at least one point in each connected component of a
 smooth real hypersurface, taking into account bit complexity and
 all coefficient growth.  This problem is a basic and important
 subroutine used in semi-algebraic geometry. For instance, it is used
 in determining an upper bound on the number of connected components
 of a real hypersurface. The algorithm requires some genericity
 properties, which we ensure through random changes of variables that
 produce generic coordinate systems. In order to ensure these
 properties with arbitrarily high probability, we developed a new
 quantitative extension of Thom's weak transversality theorem, and a
 new quantitative Noether normalization statement for polar
 varieties. These statements enable us to give a precise probability
 analysis of a monte carlo algorithm, where we can guarantee
correctness with arbitrarily high probability.
\par 
  Algorithms for deciding connectivity queries in real algebraic
  geometry have been developed in a series of papers
  \cite{a,b,c,d}. However, these algorithms are based in the
  arithmetic complexity model, and analyses of the bit complexity
  questions were left for future work.  Our results are an important
  first step in this analysis.
\end{abstract}

@@

\nl Use \cite{AD} to compute the partial derivatives $\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}$\;

@@

Let $\A$ be an $n \times n$ matrix of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define $f^{\A} \in \mathbb{Q}(\A)[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$ Recall from Section 4.1.4 that $f^\A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2),$ and $\textbf{H}_i^{'}(3).$ Throughout this section, for simplicity let $\I^\A$ denote $\I(\pi_i,V^\A)$ and let $W^\A$ denote $W(\pi_i,V^\A).$ To simplify notation, we will let $i \in \{1,\hdots,\}$ be fixed, and we wil avoid writing $i$ where the context is clear. 

@@

Let $f \in \mathbb{Q}[X_1,\hdots,X_n]$ be squarefree with total degree $D$ and $V(f)$ smooth. Recall $\textbf{H}_i^{'}(4):$ \textit{$W(\pi_i,V)$ is in Noether position for $\pi_{i-1}$}, which is equivalent to having the extension  

@@

Section 8 is devoted to proving Theorem 8.1. The majority of work is in upper bounding the degree of $\Delta_3,$ as in \cite{EMP} it was shown that $f^A$ satisfies $\textbf{H}_i^{'}$ for a generic change of variables $A\in\C^{n^2}.$ 

@@

then $A$ cancels none of the denominators of $a_{j}$ or $b_{j,l}$. Define $h_j \in \Q[\A]$ to be the product of the denominators of the coefficients of $a_j$ and $b_{j,l}$. Hence, if $A \not \in V(h_j)$ then $A$ cancels none of the denominators $a_{j}$ or $b_{j,l}$ which are providing the integral dependence relation. Define

@@

\Xi_i := V(h_i,\hdots,h_n).

@@

For each $j \in \{i,\hdots,n\}$, we want a bound on the degree of $h_j$. Then by applying Bezout's theorem we can obtain a bound on the degrees of $\Xi_i$. We can obtain bounds for $h_j$ by applying an effective Nullstellensatz theorem. However, the theorem requires $P_{j}$ to be a polynomial. We can obtain a polynomial from $P_{j}$ by clearing denominators. We can multiply $P_{j}$ by $d_j \in \Q[\A_{i,j}]$ equal to the least common multiple of the denominators of $P_{j}.$ Put 

@@

\overline{P_{j}} := d_jP_{j}\in \Q[\A][\textit{\textbf{X}}],

@@

$\deg \pjb\leq (D+1)^{n-i+1}.$ (The same for all $j$?)

@@

\deg d_j \leq \deg \pjb \leq (D+1)^{n-i+1}.

@@

\alpha_j = \sum_{l=1}^{n-i+1} A_{j,l}(\A,\textit{\textbf{X}},1/\pjb)g_l + 0

@@

A_{j,l}(\A,\textit{\textbf{X}},1/\pjb) 
&= \sum_{i=1}^{n_l} A(\A,\Xb,1/\pjb)\\
&= \sum_{i=1}^{n_l} \left(\frac{1}{P_j^{k_l}}\right)\At_{j,l,i}(\A,\textit{\textbf{X}}), ~k_l \in \mathbb{Z}.

@@

Now let $K := \max\{k_l\}$ so that 

@@

    \pjb^K \alpha_j &= \pjb^K\left(\sum_{l=1}^{n-i+1} A_{j,l}(\A,\textit{\textbf{X}},1/\pjb)g_l\right)\\
    &= \sum_{l=1}^{n-i+1}\left( \sum_{i=1}^{n_l}\left(\pjb^{K-k_{n_l}}\right)\At_{j,l,i}(\A,\Xb) \right)g_l \in \Q[\A][\Xb],

@@

which shows that $\alpha_j$ is the only remaining denominator (i.e.\ the only denominators are $d_j$ and $\alpha_j$). We know that $\alpha_j$ is not zero, we know the leading coefficient of $\overline{P_j}$ is $d_j$ which is not zero, we have an upper bound for the degrees of $\alpha_j$, and we know $\deg d \leq \deg P_j = (D+1)^{n-i+1}$, where $\pjb=d_jP_j.$ By making $\pjb$ monic (this is just $P_j$?), the integrality of 
\mathbb{Q}(\frak{A}_{i,j})[\textbf{X}_{\leq i-1}]\rightarrow\mathbb{Q}(\frak{A}_{i,j})[\textbf{X}]/\I^{\A}

@@

holds with $\pjb(X_j)\in \I^{\A}.$

@@
